Robust Estimation, Prediction and Control with \\ Linear Dynamics and Generic Costs

### 2020-02-04 main.tex
Update proof

However, under an additional smoothness assumption on $R$, we can bound this gap:
\begin{theorem}
\label{thm:control-error}
If the reward function $R$ is lipschitz, then the approximation error of \autoref{alg:full} is upper-bounded by:
\begin{equation*}
     V^r(\bu) - \hat{V}^r(\bu) = \cO\left(\frac{1}{\sqrt{N}}\right)
\end{equation*}
\end{theorem}

@@

We first bound the model estimation error.
\begin{lemma}
Under sufficient excitations in the features 
\begin{align}
\label{eq:excitation}
\exists \underline{\phi},\overline{\phi}>0, n_0: \forall n\geq n_0,i,j\in[p],\nonumber\\ \underline{\phi}^2 \leq \sum_k \sum_l \phi_{n,k,i}\Sigma_{p,k,l}^{-1}\phi_{n,l,j} \leq \overline{\phi}^2,
\end{align}
then,
\[\|\theta - \theta_{[N],\lambda}\|_{2} = \cO\left( \sqrt{\frac{\log N}{N}} \right) \]
\end{lemma}
\begin{proof}
By \eqref{eq:g_n_lambda} and \eqref{eq:excitation}, we have $G_{[N],\lambda,i,j} \geq N\underline{\phi}^2 $

@@

\begin{align*}
\beta_N(\delta) &= \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
+ (\lambda d)^{1/2}S\\
&\leq \sqrt{d\log (N\overline{\phi}^2) / \lambda^{d/2}} + \cO(1)
\end{align*}

which yields the desired bound.
\end{proof}

@@

\begin{lemma}
	\todo{Which result?}
\end{lemma}
\begin{proof}
Let $\delta x = \ox - \ux$. Assuming that eventually after $n\geq n_0$ $A_0$ becomes close enough to $A(\theta)$ to always be Metlzer. Then, \eqref{eq:interval-predictor} gives the dynamics:
\begin{align*}
\dot{\delta x} = A_0\delta x + |\Delta A|(\ox^+ + \ux^-) + |D|(\overline{\omega} - \underline{\omega})
\end{align*}
where $|M| = M^+ + M^-$.

Moreover, it holds that $-\ux^+ -\ox^- \leq \delta x \leq \ox^+ + \ux^-$, which implies $|\ux^+ + \ox^-| \leq 2 |\delta x|$.

We define $V = \delta x^\transp P \delta x$, and compute its derivative:
\begin{align*}
\dot{V} ={}& X^\transp
\begin{bmatrix}
A_0^\transp P + P A_0^\transp + Q & P|\Delta A| & P|D|\\
|\Delta A|^\transp P & -\alpha I_n & 0\\
D^\transp|P| & 0 & -\rho I_p
\end{bmatrix}
X\\
& - \delta x^\transp Q \delta x + \alpha |\ux^+ + \ox^-|^2 + \rho |\overline{\omega} - \underline{\omega}|^2
\end{align*}
with $X=\begin{bmatrix}
\delta x\\ \ux^+ + \ox^- \\ \overline{\omega} - \underline{\omega}
\end{bmatrix}$. Hence, 
\begin{align*}
\dot{V} \leq {}& X^\transp
\begin{bmatrix}
A_0^\transp P + P A_0^\transp + Q + 2\alpha^2 I_p & P|\Delta A| & P|D|\\
|\Delta A|^\transp P & -\alpha I_n & 0\\
D^\transp|P| & 0 & -\rho I_p
\end{bmatrix}
X\\
& - \delta x^\transp Q \delta x + \rho |\overline{\omega} - \underline{\omega}|^2
\end{align*}

Finally, we propagate the state prediction error bound to the pessimistic rewards and surrogate objective to get our final result.
\begin{proof}
 By the inclusion property \eqref{eq:inclusion-property}, for any sequence of controls $\bu$, dynamics $A(\theta)\in C_{[N],\delta}$ and perturbations $\underline{\bw} \leq \bw \leq \overline{\bw}$, we have that $\underline{x}_n \leq x_n \leq \overline{x}_n$, which implies that $R(x_n) \leq \max_{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}  R(x)$.

 Thus, assuming $R$ is $L$-lipschitz,
 \begin{align}
     V^r(\bu) - \hat{V}^r(\bu) &\leq \sum_{n=N+1}^\infty \gamma^n \underset{{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}}{(\max - \min)} R(x)\\
     &\leq \sum_{n=N+1}^\infty \gamma^n L \left\|\underline{x}_n(\bu) - \overline{x}_n(\bu)\right\|\\
     &\leq L/(1-\gamma) \cO\left(\sqrt{\log N / N}\right)
 \end{align}
\end{proof}

>>>>

However, under an additional smoothness assumption on $R$, we can bound this gap:
\begin{theorem}[Approximation gap]
\label{thm:control-error}
If the reward function $R$ is lipschitz, then we can bound the approximation error of \autoref{alg:full}:
\begin{equation*}
\hat{V}^r(\bu) \leq {V}^r(\bu) \leq V(\bu) \leq \hat{V}^r(\bu) + \cO\left(\frac{1}{\sqrt{N}}\right)
\end{equation*}
\end{theorem}

@@

We first bound the model estimation error.
\begin{lemma}
Under sufficient excitations in the features 
\begin{align}
\label{eq:excitation}
\exists \underline{\phi},\overline{\phi}>0, n_0: \forall n\geq n_0,i,j\in[p],\nonumber\\ \underline{\phi}^2 \leq \sum_k \sum_l \phi_{n,k,i}\Sigma_{p,k,l}^{-1}\phi_{n,l,j} \leq \overline{\phi}^2,
\end{align}
then,
\[\|A(\theta) - A(\theta_{[N],\lambda})\|_2 = \cO\left( \sqrt{\frac{\log N}{N}} \right) \]
\end{lemma}
\begin{proof}
By \eqref{eq:g_n_lambda} and \eqref{eq:excitation}, we have $G_{[N],\lambda,i,j} \geq N\underline{\phi}^2 $

@@

\begin{align*}
\beta_N(\delta) &= \sqrt{2\ln \left(\frac{\det(G_{N,\lambda})^{1/2}}{\delta\det(\lambda I_d)^{1/2}}\right)}
+ (\lambda d)^{1/2}S\\
&\leq \sqrt{d\log (N\overline{\phi}^2) / \lambda^{d/2}} + \cO(1)
\end{align*}
Thus,
\[\|\theta - \theta_{[N],\lambda}\|_{2} = \cO\left( \sqrt{\frac{\log N}{N}} \right) \]

And $A(\theta)$ belongs to a linear image of this ball. By writing a the $j^{th}$ column of a matrix $M$ as $M_j$, and its coefficient $i,j$ as $M_{i,j}$,
\begin{align*}
((A(\theta)&-A(\theta_{[N],\lambda}))^\transp (A(\theta) - A(\theta_{[N],\lambda})))_{i,j}\\
%&= (\phi_i(\theta-\theta_{[N],\lambda}))^\transp \phi_j(\theta-\theta_{[N],\lambda}) \\
&= (\theta-\theta_{[N],\lambda})^\transp\phi_{i}^{\transp}\phi_j(\theta-\theta_{[N],\lambda}) \\
&\leq \lambda_{\max}(\phi_{i}^{\transp}\phi_j) \|\theta - \theta_{[N],\lambda}\|_{2} = \cO\left( \sqrt{\frac{\log N}{N}} \right) 
\end{align*}

@@

\begin{lemma}

\[\|\ox(t) - \ux(t)\| \leq C_\omega(t) +\cO\left( \sqrt{\frac{\log N}{N}} \right) \]
where $C_\omega(t)$ is the interval size $\|\ox(t) - \ux(t)\|$ obtained by running \eqref{eq:interval-predictor} with $\Delta A=0$.

\end{lemma}
\begin{proof}
Let $e = \ox - \ux$. Assuming that eventually after $n\geq N \geq N_0$, $A_0 = A(\theta_{[N],\lambda})$ becomes close enough to $A(\theta)$ to always be Metlzer. Then, \eqref{eq:interval-predictor} gives the dynamics:
\begin{align*}
\dot{e} = A_0e + |\Delta A|(\ox^+ + \ux^-) + |D|(\overline{\omega} - \underline{\omega})
\end{align*}
where $|M| = M^+ + M^-$.

Moreover, it holds that $-\ux^+ -\ox^- \leq e \leq \ox^+ + \ux^-$, which implies $|\ux^+ + \ox^-| \leq 2 |e|$.

We define $V = e^\transp P e$, which is non-negative definite provided that
$
P>0,
$ and compute its derivative:
\begin{align*}
\dot{V} ={}& X^\transp
\begin{bmatrix}
A_0^\transp P + P A_0^\transp + Q & P|\Delta A| & P|D|\\
|\Delta A|^\transp P & -\alpha I_n & 0\\
D^\transp|P| & 0 & -\rho I_p
\end{bmatrix}
X\\
& - e^\transp Q e + \alpha |\ux^+ + \ox^-|^2 + \rho |\overline{\omega} - \underline{\omega}|^2
\end{align*}
with $X=\begin{bmatrix}
e\\ \ux^+ + \ox^- \\ \overline{\omega} - \underline{\omega}
\end{bmatrix}$. Hence, 
\begin{align*}
\dot{V} \leq {}& X^\transp
\underbrace{\begin{bmatrix}
A_0^\transp P + P A_0^\transp + Q + 2\alpha^2 I_p & P|\Delta A| & P|D|\\
|\Delta A|^\transp P & -\alpha I_n & 0\\
D^\transp|P| & 0 & -\rho I_p
\end{bmatrix}}_{\Upsilon}
X\\
& - e^\transp Q e + \rho |\overline{\omega} - \underline{\omega}|^2
\end{align*}

Thus, if $\Upsilon < 0$, $Q>0$, $\rho > 0$, $V$ is bounded and can only increases from $V(e(t_N)) = 0$ to a $V(e)$ such that:
\[
e^\transp Q e = X^\transp \Upsilon X + \rho |\overline{\omega} - \underline{\omega}|^2
\]
\end{proof}

Finally, we propagate the state prediction error bound to the pessimistic rewards and surrogate objective to get our final result.
\begin{proof}

 For any sequence of controls $\bu$, dynamics $A(\theta)\in C_{[N],\delta}$ and perturbations $\underline{\bw} \leq \bw \leq \overline{\bw}$, we clearly have 
 \[V(\bu)^r \leq V(\bu) = \expectedvalue_{\bw}\sum_n \gamma^n R(x_n)\]

 Moreover, by the inclusion property \eqref{eq:inclusion-property}, we have that $\underline{x}_n \leq x_n \leq \overline{x}_n$, which implies that $R(x_n) \leq \max_{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}  R(x)$. Assuming $R$ is $L$-lipschitz,
 \begin{align}
     V(\bu) - \hat{V}^r(\bu) &\leq \sum_{n=N+1}^\infty \gamma^n \underset{{x\in[\underline{x}_n(\bu), \overline{x}_n(\bu)]}}{(\max - \min)} R(x)\\
     &\leq \sum_{n=N+1}^\infty \gamma^n L \left\|\underline{x}_n(\bu) - \overline{x}_n(\bu)\right\|_2\\
     &\leq L\sum_{n>N} \gamma^n C_{\omega,n} + \cO\left(\sqrt{\log N / N}\right))
 \end{align}
\end{proof}
